# The Policy Learning through Information Processing Model: An Application

## Introduction

The learning through information processing model focuses on belief change following the processing of information. Beliefs are thought to be of one of three types, including deep core, policy core, and secondary. Learning is understood as a change in beliefs. The model posits that belief change is a function of the strength of the prior belief and the strength of the information signal.

In this chapter, to illustrate the utility of our approach, we examine US public opinion data on nuclear energy before and after the Fukushima nuclear accident. The Fukushima accident in Japan was a focusing event that served as a strong information signal about the risks of nuclear energy. In this case, we examine how policy core beliefs, specifically the issue definition of nuclear energy, and preferences for constructing new nuclear power plants changed as a result of Fukushima.

Before turning to the data, it is important to justify our use of public opinion, rather than a sample of policy elites, to test the learning through information processing model. Our approach to belief systems draws on the ACF, which is typically applied to coalitions of elite policy actors [@sabatierPolicyChangeLearning1993; @nohrstedtAdvocacyCoalitionFramework2023]. However, our interest is in the structure of belief systems and in the process by which beliefs update in response to information, not in the coalition dynamics or any other major aspects of the ACF. Modeling how the public, rather than elites, update political beliefs and attitudes in response to information is a long-standing tradition in political behavior research, one that engages directly with information processing among the mass public rather than borrowing from elite-level theory. Formal Bayesian models of citizen learning have a substantial history in this literature [e.g., @gerberRationalLearningPartisan1998; @bullockPartisanBiasBayesian2009], and a large body of related work examines how the members of the public receive, resist, and are persuaded by political information [@zallerNatureOriginsMass1992; @zallerPoliticalAwarenessElite1990; @lodgePartisanSchemaPolitical1986; @kuklinskiReconsideringRationalPublic2000]. Our approach fits within this established literature on information processing among the public.

A second, complementary justification concerns whether the structure of belief systems, most notably whether beliefs are organized hierarchically and update coherently in response to information, is comparable between elites and the public. Foundational work on belief systems in mass publics found that constraint among political beliefs is not uniform across the public, but concentrated among the more politically sophisticated and attentive segment of citizens, who reason about political issues in ways that closely resemble elites [@converseNatureBeliefSystems1964a]. Building on this insight, subsequent research has directly compared elite and mass belief systems using the same core, policy core, and secondary belief hierarchy, and in a closely related substantive domain. @herronPerceptionsNuclearSecurity2002 compared two elite samples (physical scientists and state legislators) with two general-public samples on beliefs about nuclear security, a domain chosen because its complexity and historical secrecy make it a demanding test of whether structure can be found among the mass public at all. They found that core beliefs (political ideology) predicted policy core beliefs (perceived risks and benefits), which in turn predicted secondary preferences, in the same direction and with the same statistically significant pattern among the both public and elites. However, elite belief systems were more tightly integrated, with core and policy core beliefs explaining substantially more variance in secondary preferences among elites than among the public. In other words, the *structure* of belief systems replicated across elite and mass samples even though elite beliefs were more strongly organized. @jenkins-smithForeignDomesticPolicy2004 extend this finding across a different comparison, showing that the same hierarchical structure of nuclear-related beliefs holds among mass publics in both the United States and Great Britain, despite the two countries' publics differing substantially in their average beliefs, thus providing evidence that this hierarchical structure is not an artifact of a single national context. In a different substantive domain, @tumlisonCulturalValuesTrust2019 similarly found that the same values-to-trust-to-risk perception pathway operated among both policy elites and the general public, with one notable difference in degree as trust in information sources played a larger mediating role for elites than for the public. Taken together, this work finds that the *relationships* among beliefs, specifically the hierarchical structure, replicate across elite and mass samples and across national contexts, even as elite belief systems remain somewhat more tightly constrained than those of the public. Following this logic, rather than relying on the public as a whole, we restrict our analytic sample to respondents with an advanced degree beyond a bachelor's, a common strategy for approximating the more attentive and sophisticated element of the public [as did @ripbergerCulturalTheoryMeasurement2014]. Additionally, public opinion data has been used more broadly to study belief systems and policy learning [@jenkins-smithRockHardPlace2009; @ripbergerCulturalTheoryMeasurement2014; @nowlinPolicyLearningInformation2021; @offer-westortWhichFrameFits2026; @nowlinLearningUncertaintyInformation2025; @nowlinPolicyBeliefsBelief2024].

A substantial body of research questions whether citizens update political beliefs in a considered, Bayesian fashion at all, arguing instead that political reasoning is often driven by partisan or group identity rather than by the evidence at hand [@achenDemocracyRealistsWhy2017a]. This critique is aimed principally at the low-information, low-attentiveness citizens who make up the bulk of the mass public, which are the respondents excluded by our restriction to an advanced-degree sample. Also, related work defends Bayesian learning models of citizen attitudes once measurement error and motivated reasoning are properly accounted for [@gerberRationalLearningPartisan1998]. In addition, the Fukushima accident is the kind of highly salient, widely covered focusing event that research on low-information rationality suggests citizens can process meaningfully even without specialized issue expertise [@lupiaDemocraticDilemmaCan1998; @kuklinskiCitizenKnowledgeChoices1982], further supporting its use as a strong information signal for the public as well as for elites. Finally, public opinion data offers the practical advantage of a much larger number of respondents and greater variation than would be available in a sample of policy elites; for this project, this allows us to draw on 17 years of data regarding beliefs and preferences regarding nuclear energy. Taken together, these considerations support public opinion, and particularly the advanced-degree subsample used here, as a useful population in which to test the learning through information processing model.

## Nuclear Energy Beliefs and Preferences 

In March 2011, a large tsunami hit the coast of Japan. The nuclear power plant in Fukushima was flooded, which led to a loss of power in the plant and a subsequent release of radioactive material. Like the two previous nuclear accidents, Three Mile Island and Chernobyl, the accident at the nuclear power plant at Fukushima provided a potentially strong information signal about the risks of nuclear energy, and public support for nuclear energy tends to fall in the wake of such accidents [@besleyImpactAccidentAttention2014; @guptaTrackingNuclearMood2019]. Additionally, as we discuss in Chapter 4, many countries began to reconsider their approach to nuclear energy [@aldrichTriggersPolicyChange2019; @daviesFukushimasShadow2015; @namComparativeAnalysisDecision2021]. Focusing events, such as accidents at nuclear power plants, can make it more likely that those issues reach the policymaking agenda [@birklandFocusingEventsMobilization1998]; however, focusing events and the information signals they send also create opportunities for policy learning.

In the context of the information processing and learning model, the Fukushima accident is a strong information signal about nuclear energy, likely to attract the attention of policy actors, including the public. Yet, the signal is filtered through their belief systems, and, as laid out in the model of learning through information processing, learning occurs when beliefs are updated in the direction of the information. As noted, policy actors are assumed to have a three-tiered belief system that includes deep core beliefs, policy core beliefs, and secondary beliefs. Within the belief hierarchy, beliefs move from abstract foundational beliefs -- deep core beliefs -- to beliefs specific to a policy issue or subsystem -- policy core beliefs -- to secondary beliefs, which are about specific means to achieve policy goals. Deep core beliefs inform and constrain policy core beliefs, and both deep core and policy core shape secondary beliefs. The model of learning through information processing focuses on the updating of beliefs resulting from the processing of policy-relevant information. While learning as belief updating could occur for any type of belief, as beliefs become increasingly applicable to multiple policy issues, they become less likely to change. It is rare that deep core beliefs, such as political ideology or cultural worldviews, change, whereas narrowly focused secondary beliefs are the most likely to change. Policy core beliefs are more likely to change than deep core beliefs, but less likely than secondary aspects. One way for learning to occur regarding policy core beliefs is through a shift in the issue definition that undergirds policy core beliefs.

At the heart of policy core beliefs is the issue definition, or the various dimensions of the issue weighted by their salience. As dimensions are added or the salience of the dimensions change, learning regarding policy core beliefs becomes more likely. Following changes to the issue definition, policy core beliefs are likely to be updated. With the model of policy learning through information processing, learning occurs when the information signal is strong relative to the prior beliefs. Strongly held prior beliefs would only change with an extremely strong and heavily weighted information signal. In the case of nuclear energy, the Fukushima accident was a strong signal about the risks of nuclear energy and likely made the dimensions of risks associated with nuclear energy more salient, thus altering the issue definition and thereby altering the policy core beliefs. As shown in @fig-defs, policy preferences may change following the updating of policy core beliefs, and @fig-model illustrates the process for nuclear energy.

```{r}
#| label: fig-model
#| fig-cap: "Ideology, Nuclear Energy Beliefs, and Preferences for More Nuclear Plants"
#| echo: false
#| message: false
#| out-width: "100%"

library(DiagrammeR)
library(DiagrammeRsvg)
library(rsvg)

model_graph <- grViz("
digraph model {
  graph [rankdir = LR, fontname = Helvetica]
  node [shape = box, style = filled, fillcolor = white, fontname = Helvetica, fontsize = 12]
  edge [fontname = Helvetica, fontsize = 10]

  A [label = 'Political Ideology']
  subgraph cluster_id {
    label = 'Issue Definition'
    fontname = 'Helvetica-Bold'
    B [label = 'Benefits']
    C [label = 'Risks']
    B -> C [dir = both]
  }
  D [label = 'More Nuclear Plants']

  A -> B [label = 'Information\nProcessing', fontname = 'Helvetica-Oblique']
  C -> D [label = 'Information\nProcessing', fontname = 'Helvetica-Oblique']
}
")

model_svg <- export_svg(model_graph)
dir.create('figures', showWarnings = FALSE)
model_png <- 'figures/fig-model-generated.png'
rsvg_png(charToRaw(model_svg), file = model_png, width = 1600)
knitr::include_graphics(model_png)
```

As shown in @fig-model, information is processed relative to deep core beliefs (i.e., political ideology), and as a result, the issue definition of nuclear energy (i.e., policy core beliefs) is potentially updated. Note that the issue definition of nuclear energy is constituted by dimensions associated with the benefits and risks of nuclear energy. The updating of policy preferences results from the same learning dynamic. Information processing occurs along the chain of deep core beliefs $\rightarrow$ policy core beliefs $\rightarrow$ policy preferences, and updating occurs when the information signal is sufficiently strong relative to the prior belief or preference. Learning happens when the beliefs or the certainty associated with the beliefs change significantly. To estimate the model shown in @fig-model, we use survey data from an elite sample, those with post-bachelors degrees, of the US public collected over 17 years.

## Data and Measures

```{r load data and packages}
#| echo: false
#| include: false
#| warning: false

## load packages 
library(QuantPsyc)

## load data 
nukeData <- read.csv("data/nukeData.csv")

## recodes 
nukeData$nukeRisk <- round((nukeData$nrsk1+nukeData$nrsk2+nukeData$nrsk3+
                   nukeData$nrsk4)/4,0)
nukeData$nukeBenefit <- round((nukeData$nben1+nukeData$nben2+nukeData$nben3+
                               nukeData$nben4)/4,0)
nukeRiskA <- psy::cronbach(data.frame(nukeData$nrsk1,nukeData$nrsk2,nukeData$nrsk3,
                                      nukeData$nrsk4))
nukeBenefitA <- psy::cronbach(data.frame(nukeData$nben1,nukeData$nben2,nukeData$nben3,
                                      nukeData$nben4))
nukeData$fuk <- ifelse(nukeData$year==2012 | 
                           nukeData$year==2013 |
                           nukeData$year==2014 |
                           nukeData$year==2015 |
                           nukeData$year==2016 |
                           nukeData$year==2017 |
                           nukeData$year==2018 |
                           nukeData$year==2019 |
                           nukeData$year==2020 |
                           nukeData$year==2021 |
                           nukeData$year==2022 |
                           nukeData$year==2023,1,0)


nukeDataC <- subset(nukeData, college==1) # subset of college educated
nukeDataC2 <- subset(nukeData, grad==1) # subset of advanced degree

nukeRiskA <- psy::cronbach(data.frame(nukeDataC2$nrsk1,nukeDataC2$nrsk2,nukeDataC2$nrsk3,
                                      nukeDataC2$nrsk4))
nukeBenefitA <- psy::cronbach(data.frame(nukeDataC2$nben1,nukeDataC2$nben2,nukeDataC2$nben3,
                                      nukeDataC2$nben4))

preData <- as.data.frame(subset(nukeDataC2, fuk==0))
postData <- as.data.frame(subset(nukeDataC2, fuk==1))
```

To examine the impact of the Fukushima accident on beliefs about nuclear energy, we draw on data from 17 years, 2006-2023, of nationwide surveys of US residents.[^chap4-1] The Institute for Public Policy Research and Analysis at the University of Oklahoma developed and administered the surveys, which were fielded annually. The quota-based samples for each survey were recruited by firms such as Survey Sampling International and Qualtrics, and the quotas for gender, race, and age helped to ensure that the samples were as representative as possible of the US population.[^chap4-2] In each survey, respondents were asked about their views on energy and environmental issues, including nuclear energy. Additionally, the surveys gauged respondents' deep core beliefs by including measures of political ideology. As discussed above, we restrict our analytic sample to respondents who have completed an advanced degree, in order to approximate the more attentive and sophisticated stratum of the public in which elite-like belief structures are most likely to be present. Overall, the combined surveys include `r nrow(nukeDataC2)` respondents with advanced degrees, which we use for the subsequent analysis. Next, we discuss the measures we used to examine beliefs and policy preferences regarding nuclear energy.^[The descriptive statistics and pairwise correlations for the variables used throughout this chapter's analysis are reported in the appendix to this chapter.]

[^chap4-1]: Some years are missing due to question wording changes. We thank Dr. Hank Jenkins-Smith and Dr. Kuhika Gupta for the use of the data.

[^chap4-2]: Response rates for these surveys are not available, which is a limitation of the data, since response rate is one indicator commonly associated with the risk of non-response bias. However, response rate alone is an incomplete measure of that risk, because what matters for bias is whether respondents differ systematically from non-respondents on the characteristics of interest, not the rate of response itself [@grovesTotalSurveyError2010]. The quota-based sampling approach used here, which matches the sample to US Census-based population data on gender, race, and age, is a widely used method for improving the representativeness of non-probability samples such as ours [@bakerResearchSynthesisAAPOR2010]. 

### Deep Core Beliefs

For deep core beliefs, we use a measure of self-reported political ideology. Among the US public, support for nuclear energy is polarized along ideological lines, with conservatives, on average, more supportive of nuclear energy than liberals [@hawesClimateSciencePolitics2022; @kuklinskiCitizenKnowledgeChoices1982; @besleyImpactAccidentAttention2014; @mcbethPoliticalIdeologyNuclear2023]. To measure political ideology, respondents were asked, "On a scale of political ideology, individuals can be arranged from strongly liberal to strongly conservative. Which of the following categories best describes your views?", with 1 indicating strongly liberal, 4 = middle of the road, and 7 = strongly conservative. The mean on the political ideology scale is `r round(mean(nukeDataC2$ideology, na.rm = TRUE),2)`, which is about middle of the road.

### Policy Core Beliefs: The Dimensions of Nuclear Energy

To explore the issue dimensions of nuclear energy, we rely on a set of questions about its risks and benefits. The assumption is that views about the risks and benefits are the key drivers of how the nuclear energy issue is defined, and that the weighting of those dimensions shapes the policy core beliefs about nuclear energy.

Respondents were asked to rate, on a zero to ten scale, their perceived risks of nuclear energy across four dimensions of risk, including an accident at a nuclear power plant, an accident during transportation or storage of nuclear materials, a terrorist attack on a nuclear plant, and diversion of fuel to make a nuclear weapon. The exact question wording was as follows:

> First we want to know about your beliefs concerning some of the possible risks associated with nuclear energy use in the U.S. Please consider both the likelihood of a nuclear event occurring and its potential consequences when evaluating the risk posed by each of the following on a scale from zero to ten, where zero means no risk and ten means extreme risk. \[*Statements were presented in random order*\]

- **Plant Accident**: An event at a U.S. nuclear power plant within the next 20 years that results in the release of large amounts of radiation

- **Transportation Accident**: An event during the transportation and temporary storage of spent nuclear fuel from nuclear power plants in the U.S. within the next 20 years that results in the release of large amounts of radiation

- **Terrorist Attack**: A terrorist attack at a U.S. nuclear power plant within the next 20 years that results in the release of large amounts of radiation

- **Weapon**: The diversion of nuclear fuel from a nuclear power plant in the U.S. within the next 20 years for the purpose of building a nuclear weapon

Similarly, respondents were asked about the benefits of nuclear energy across four dimensions, including no greenhouse gas emissions, a reliable source of power, contributions to US energy independence, and reduced environmental damage compared to coal or oil and gas extraction.

> Next we want to know about your beliefs concerning some of the possible benefits associated with nuclear energy use in the U.S. Please evaluate each of these possible benefits of nuclear energy use on a scale from zero to ten, where zero means not at all beneficial and ten means extremely beneficial. \[*Statements were presented in random order*\]

- **No GHG**: Reducing environmental threats because the generation of nuclear energy does not produce greenhouse gases that are believed to cause climate change

- **Reliable Power**: Reliable power because nuclear energy generates large amounts of electricity and is not affected by weather conditions, such as low rainfall or no wind

- **Energy Independence**: Greater U.S. energy independence because nuclear energy production does not require oil or gas from foreign sources

- **Reduced Environmental Damage**: Reduced environmental damage because of less need for mining coal or extracting oil and gas

Additionally, we combined and averaged the risk and benefit items into two scales. The Cronbach's $\alpha$ for the risk items is `r round(nukeRiskA$alpha,3)`, and the $\alpha$ for the benefit items was `r round(nukeBenefitA$alpha,3)`.

### Policy Preferences

To measure policy preferences, we examine support for building more nuclear power plants in the US. Specifically, we averaged responses to two questions: one on support for adding additional reactors at existing nuclear power plants and the other on support for adding new reactors at new locations in the US.[^chap4-3]

[^chap4-3]: We do not have a measure of secondary beliefs, rather we assume that secondary beliefs are implied by policy preferences.

> Using a scale from one to seven, where one means *strongly oppose* and seven means *strongly support*, how do you feel about constructing additional nuclear reactors at the sites of existing nuclear power plants in the U.S.?

> Using the same scale from one to seven, where one means *strongly oppose* and seven means *strongly support*, how do you feel about constructing additional nuclear power plants at new locations in the U.S.?

For analysis, we combine and average the two scales into a single one to seven measure of support for more nuclear energy plants. The average support for more nuclear plants is `r round(mean(nukeDataC2$morePlants, na.rm = TRUE),3)`, which is above the scale midpoint of 4.

## Results

```{r diff of means}
#| echo: false
#| include: false
#| warning: false

mean(preData$nrsk1, na.rm = TRUE)
mean(postData$nrsk1, na.rm = TRUE)
t.test(preData$nrsk1,postData$nrsk1)

mean(preData$nrsk2, na.rm = TRUE)
mean(postData$nrsk2, na.rm = TRUE)
t.test(preData$nrsk2,postData$nrsk2)

mean(preData$nrsk3, na.rm = TRUE)
mean(postData$nrsk3, na.rm = TRUE)
t.test(preData$nrsk3,postData$nrsk3)

mean(preData$nrsk4, na.rm = TRUE)
mean(postData$nrsk4, na.rm = TRUE)
t.test(preData$nrsk4,postData$nrsk4)

mean(preData$nukeRisk, na.rm = TRUE)
mean(postData$nukeRisk, na.rm = TRUE)
t.test(preData$nukeRisk,postData$nukeRisk)

mean(preData$nben1, na.rm = TRUE)
mean(postData$nben1, na.rm = TRUE)
t.test(preData$nben1,postData$nben1)

mean(preData$nben2, na.rm = TRUE)
mean(postData$nben2, na.rm = TRUE)
t.test(preData$nben2,postData$nben2)

mean(preData$nben3, na.rm = TRUE)
mean(postData$nben3, na.rm = TRUE)
t.test(preData$nben3,postData$nben3)

mean(preData$nben4, na.rm = TRUE)
mean(postData$nben4, na.rm = TRUE)
t.test(preData$nben4,postData$nben4)

mean(preData$nukeBenefit, na.rm = TRUE)
mean(postData$nukeBenefit, na.rm = TRUE)
t.test(preData$nukeBenefit,postData$nukeBenefit)

t.test(preData$nukeBenefit,preData$nukeRisk)
t.test(postData$nukeBenefit,postData$nukeRisk)
```

```{r models}
#| echo: false
#| include: false
#| warning: false

## ols models 
summary(olsRiskFuk <- lm(nukeRisk~age+gender+white+ideology+nukeBenefit+fuk, data = subset(nukeData, grad==1)))
summary(olsBenefitFuk <- lm(nukeBenefit~age+gender+white+ideology+nukeRisk+fuk, data = subset(nukeData, grad==1)))
summary(olsMorePlantsFuk <- lm(morePlants~nukeRisk+nukeBenefit+age+gender+white+ideology+fuk, 
                            data = subset(nukeData, grad==1)))

## standardized betas 
lm.beta(olsRiskFuk)
lm.beta(olsBenefitFuk)
lm.beta(olsMorePlantsFuk)
```

```{r}
#| echo: false
#| warning: false
#| label: fig-rbGraph
#| fig.width: 8
#| fig.height: 5
#| fig-cap: Aggregate Perceptions of the Risks and Benefits of Nuclear Energy, 2006-2023  

library(ggplot2) 
library(dplyr)

riskBenefitYear <- nukeDataC2 %>% 
    group_by(year) %>% 
    summarise(nukeRisk = mean(nukeRisk, na.rm = TRUE),
              nukeBenefit = mean(nukeBenefit, na.rm = TRUE))  

ggplot(data=riskBenefitYear, aes(x=year)) +
  geom_line(aes(y=nukeRisk), linetype=2) +
  geom_line(aes(y=nukeBenefit)) +
  geom_point(aes(y=nukeRisk)) +
  geom_point(aes(y=nukeBenefit)) + 
  theme_classic() +
  labs(x="Year",
       y="Perceived Risk/Benefit Scales") +
  ylim(0,10) +
  geom_vline(xintercept = 2011, linetype=3) +
  geom_text(x=2012.25, y=0, label="Fukushima", fontsize=10, fontface="italic") +
  geom_text(x=2006, y=5, label="Risk", fontsize=10) +
  geom_text(x=2006, y=8, label="Benefit", fontsize=10)


```

We first plot the means of the combined risk and benefit scales for each year, as shown in @fig-rbGraph. As can be seen, the average perception of risks increased following the Fukushima accident, and continued to increase. However, on average, perceptions of benefits decreased sightly, but remained relatively consistent over time.

Next, we examine differences in means of the dimensions of nuclear energy -- the measure of risks and benefits -- pre- and post the Fukushima accident. The pre-Fukushima measures include pooled survey data from 2006 to 2011, and post-Fukushima is pooled survey data from 2012 to 2023. @tbl-dims shows the pre- and post-Fukushima means for each dimension and the averaged items. The table also includes the difference in means for each dimension, and indicates whether those differences are statistically significant. 

| **Scale** | **Issue Dimensions** | **Pre-Fukushima Mean** | **Post-Fukushima Mean** | **Difference** |
|:--------------|:--------------|-------------:|-------------:|-------------:|
| **Risks (0 to 10 scale)** | Plant Accident<br> | `r round(mean(preData$nrsk1, na.rm = TRUE),2)` | `r round(mean(postData$nrsk1, na.rm = TRUE),2)` | `r round(mean(postData$nrsk1, na.rm = TRUE),2)-round(mean(preData$nrsk1, na.rm = TRUE),2)`\*\*\* |
|  | Transportation Accident<br> | `r round(mean(preData$nrsk2, na.rm=TRUE),2)` | `r round(mean(postData$nrsk2, na.rm=TRUE),2)` | `r round(mean(postData$nrsk2, na.rm=TRUE),2)-round(mean(preData$nrsk2, na.rm=TRUE),2)`\*\*\* |
|  | Terrorist Attack<br> | `r round(mean(preData$nrsk3, na.rm=TRUE),2)` | `r round(mean(postData$nrsk3, na.rm=TRUE),2)` | `r round(mean(postData$nrsk3, na.rm=TRUE),2)-round(mean(preData$nrsk3, na.rm=TRUE),2)`\*\*\* |
|  | Weapon<br> | `r round(mean(preData$nrsk4, na.rm=TRUE),2)` | `r round(mean(postData$nrsk4, na.rm=TRUE),2)` | `r round(mean(postData$nrsk4, na.rm=TRUE),2)-round(mean(preData$nrsk4, na.rm=TRUE),2)`\*\*\* |
|  | **All Items** | `r round(mean(preData$nukeRisk, na.rm=TRUE),2)` | `r round(mean(postData$nukeRisk, na.rm=TRUE),2)` | `r round(mean(postData$nukeRisk, na.rm=TRUE),2)-round(mean(preData$nukeRisk, na.rm=TRUE),2)`\*\*\* |
| **Benefits (0 to 10 scale)** | No GHG | `r round(mean(preData$nben1, na.rm=TRUE),2)` | `r round(mean(postData$nben1, na.rm = TRUE),2)` | `r round(mean(postData$nben1, na.rm = TRUE),2)-round(mean(preData$nben1, na.rm=TRUE),2)`\*\*\* |
|  | Reliable Power<br> | `r round(mean(preData$nben2, na.rm=TRUE),2)` | `r round(mean(postData$nben2, na.rm = TRUE),2)` | `r round(mean(postData$nben2, na.rm = TRUE),2)-round(mean(preData$nben2, na.rm=TRUE),2)` |
|  | Energy Independence<br> | `r round(mean(preData$nben3, na.rm=TRUE),2)` | `r round(mean(postData$nben3, na.rm = TRUE),2)` | `r round(mean(postData$nben3, na.rm = TRUE),2)-round(mean(preData$nben3, na.rm=TRUE),2)`\*\* |
|  | Reduced Environmental Damage<br> | `r round(mean(preData$nben4, na.rm=TRUE),2)` | `r round(mean(postData$nben4, na.rm=TRUE),2)` | `r round(mean(postData$nben4, na.rm=TRUE),2)-round(mean(preData$nben4, na.rm=TRUE),2)` |
|  | **All Items** | `r round(mean(preData$nukeBenefit, na.rm=TRUE),2)` | `r round(mean(postData$nukeBenefit, na.rm=TRUE),2)` | `r round(mean(postData$nukeBenefit, na.rm=TRUE),2)-round(mean(preData$nukeBenefit, na.rm=TRUE),2)`\* |

: Comparison of Nuclear Energy Issue Dimensions Pre-and-Post-Fukushima {#tbl-dims}

*Note:* \* p\<0.05; \*\* p\<0.01; \*\*\* p\<0.001

As shown in @tbl-dims and consistent with @fig-rbGraph, the mean of each risk measure increased significantly following Fukushima. Additionally, the mean of the scale combining the risk measures increased. For the benefit measures, *no greenhouse gases* and *energy independence* were significantly lower on average after Fukushima; however, there were no significant differences in the *reliable power* and *reduced environmental damage* measures. Finally, the difference between the pre- and post-Fukushima combined benefits scale was significant, with the benefits of nuclear energy rated lower following Fukushima.

Also of note is that the average of the benefit measures had a significantly higher mean than the average of the risk measures both pre- and post-Fukushima. Pre-Fukushima, the average of the risk measures was `r round(mean(preData$nukeRisk, na.rm = TRUE),3)`, and the average of the benefit measures was `r round(mean(preData$nukeBenefit, na.rm = TRUE),3)`, and the difference was statistically significant \[*t* = `r round(t.test(preData$nukeRisk,preData$nukeBenefit)$statistic[["t"]],2)`, *p*\<0.001\]. Following Fukushima, the mean of the risk measures was `r round(mean(postData$nukeRisk, na.rm = TRUE),3)`, the mean of the benefit measures was `r round(mean(postData$nukeBenefit, na.rm = TRUE),3)`, and the difference was significant \[*t* = `r round(t.test(postData$nukeRisk,postData$nukeBenefit)$statistic[["t"]],2)`, *p*\<0.001\].

Next, we conduct an OLS regression analysis to examine the influence of political ideology and the Fukushima accident on the issue definition of nuclear energy, specifically the perceived risks and benefits. Then, we explore the impacts of ideology, Fukushima, and the risks and benefits of nuclear energy on support for more nuclear power plants. Note that we also control for demographics, including age, gender (male=1), and race (white=1). The results are shown in @tbl-results.

```{r}
#| echo: false
#| warning: false
#| label: tbl-results
#| tbl-cap: OLS Analysis of Political Ideology, Fukushima, Nuclear Energy Risks and Benefits, and Support for More Plants 

library("tidyverse")
library("modelsummary")

gm <- tibble::tribble(
  ~raw,        ~clean,          ~fmt,
  "nobs",      "N",             0,
  "r.squared", "R<sup>2</sup>", 2)
cm <- c("ideology" = "Ideology (more conservative)",
        "fuk" = "Fukushima",
        "nukeBenefit" = "Benefits", 
        "nukeRisk" = "Risks",
        "age" = "Age",
        "gender" = "Gender (male=1)",
        "white"= "Race (white=1)")

modelsTotal <- list("Risks"=olsRiskFuk, "Benefits"=olsBenefitFuk, "More Plants"=olsMorePlantsFuk)
modelsummary(modelsTotal, stars = TRUE, coef_map = cm, gof_map = gm)
```

*Note:* Standard errors appear in parentheses below each coefficient.

As shown in @tbl-results, increasing conservatism is associated with lower risk perceptions, higher perceptions of benefits, and greater support for more nuclear plants. In addition, perceptions of risks increased and perceptions of benefits, and support for more plants both decreased following the Fukushima accident, on average. Interestingly, the benefits and risks of nuclear energy are positively associated at *p*\<0.10. But, as expected, an increase in the perceived risks is negatively associated with support for more plants, whereas an increase in perceived benefits is positively associated. For the demographic variables, age is negatively associated with perceived risks, perceived benefits, and support for more plants. Finally, both males and whites perceive less risk, greater benefits, and higher levels of support for more nuclear power plants than non-males and non-whites.

Our final analysis is a path analysis to show the influence of political ideology, Fukushima, and beliefs about the risks and benefits of nuclear energy on preferences for more nuclear power plants.[^chap4-4] Path analysis includes standardized coefficients, which allow us to compare the impact of the various independent variables on the dependent variable. In addition, path analysis accounts for the various effects of political ideology and Fukushima on preferences for more nuclear power plants. @fig-results shows the results of the path analysis, including the standardized betas.

[^chap4-4]: For the path analysis we used the standardized coefficients from the OLS regression shown in @tbl-results.

```{r}
#| label: fig-results
#| fig-cap: "Path Analysis of Ideology, Fukushima, Nuclear Energy Beliefs, and Preferences for More Nuclear Plants"
#| echo: false
#| message: false
#| out-width: "100%"

library(DiagrammeR)
library(DiagrammeRsvg)
library(rsvg)

get_beta_stars <- function(model, var) {
  b <- unname(round(lm.beta(model)[var], 3))
  p <- summary(model)$coefficients[var, 4]
  stars <- if (p < 0.001) "***" else if (p < 0.01) "**" else if (p < 0.05) "*" else ""
  paste0(sprintf("%.3f", b), stars)
}

lbl_A_B <- get_beta_stars(olsBenefitFuk, "ideology")
lbl_A_C <- get_beta_stars(olsRiskFuk, "ideology")
lbl_A_D <- get_beta_stars(olsMorePlantsFuk, "ideology")
lbl_E_B <- get_beta_stars(olsBenefitFuk, "fuk")
lbl_E_C <- get_beta_stars(olsRiskFuk, "fuk")
lbl_B_D <- get_beta_stars(olsMorePlantsFuk, "nukeBenefit")
lbl_C_D <- get_beta_stars(olsMorePlantsFuk, "nukeRisk")
lbl_E_D <- get_beta_stars(olsMorePlantsFuk, "fuk")

results_graph <- grViz(sprintf("
digraph results {
  graph [rankdir = LR, fontname = Helvetica]
  node [shape = box, style = filled, fillcolor = white, fontname = Helvetica, fontsize = 12]
  edge [fontname = Helvetica, fontsize = 10]

  A [label = 'Political Ideology\n(more conservative)']
  E [label = 'Fukushima']
  B [label = 'Benefits']
  C [label = 'Risks']
  D [label = 'More Nuclear Plants']

  A -> B [label = '%s']
  A -> C [label = '%s']
  A -> D [label = '%s']
  E -> B [label = '%s']
  E -> C [label = '%s']
  B -> D [label = '%s']
  C -> D [label = '%s']
  E -> D [label = '%s']
}
", lbl_A_B, lbl_A_C, lbl_A_D, lbl_E_B, lbl_E_C, lbl_B_D, lbl_C_D, lbl_E_D))

results_svg <- export_svg(results_graph)
dir.create('figures', showWarnings = FALSE)
results_png <- 'figures/fig-results-generated.png'
rsvg_png(charToRaw(results_svg), file = results_png, width = 1600)
knitr::include_graphics(results_png)
```

*Note:* \* p\<0.05; \*\* p\<0.01; \*\*\* p\<0.001

As shown in @fig-results, increasing conservatism is associated with an increase in the perceived benefits of nuclear energy, a decrease in the perceived risks associated with nuclear energy, and an increase in the support for more nuclear power plants. Additionally, the Fukushima accident is associated with a decrease in perceived benefits, an increase in perceived risks, and a decrease in support for more nuclear plants. However, comparing the standardized coefficients, we see that ideology had a larger influence on each outcome measure than Fukushima, with the perceived benefits coefficient for ideology at `r round(lm.beta(olsBenefitFuk)[4],3)` and the coefficient for Fukushima is `r round(lm.beta(olsBenefitFuk)[6],3)`. Similarly, the perceived risk estimates are `r round(lm.beta(olsRiskFuk)[4],3)` for ideology and `r round(lm.beta(olsRiskFuk)[6],3)` for Fukushima.

Looking at the measure of support for more nuclear plants, the policy core beliefs regarding the perceived risks and benefits have the most influence, with the perceived benefits having a greater impact than perceived risks at `r round(lm.beta(olsMorePlantsFuk)[2],3)` for benefits versus `r round(lm.beta(olsMorePlantsFuk)[1],3)` for risks. Finally, the coefficients for ideology and Fukushima were `r round(lm.beta(olsMorePlantsFuk)[6],3)` and `r round(lm.beta(olsMorePlantsFuk)[7],3)`, respectively, indicating that ideology had a stronger effect on supporting more nuclear power plants than the Fukushima accident.

Path analysis also allows us to examine the effect of political ideology on support for more nuclear plants (*direct effects*), the mediation effect of political ideology operating through the issue definition (*indirect effects*), and the combined effects of political ideology and the issue definition of nuclear energy on support for more nuclear plants (*total effects*). The indirect effect is calculated by multiplying the path (a) from political ideology to the issue definition and the path (b) from the issue definition to support for more plants. Below, we illustrate this process for the effects of ideology and risk perceptions on support for more nuclear power plants.

- The first, simplified, model shows ideology and more plants: $\text{more plants}{_i} = \beta_0 + \beta_1(\text{ideology}{_i}) + \epsilon_i$
- Then, ideology and risk: $\text{risk}{_i} = \beta_0 + \beta_2(\text{ideology}{_i}) + \epsilon_i$
- Finally, ideology, risk, and more plants: $\text{more plants}{_i} = \beta_0 + \beta_3(\text{ideology}{_i}) + \beta_4(\text{risk}{_i}) +  \epsilon_i$

Using the standardized coefficients shown in @fig-results, we can calculate the indirect, direct, and total effects. The direct effect, $\beta_3$, is `r round(lm.beta(olsMorePlantsFuk)[6],3)`. The indirect effect is $\beta_2 * \beta_4$, `r round(lm.beta(olsRiskFuk)[4]*lm.beta(olsMorePlantsFuk)[1],3)`. Finally, the total effect, $(\beta_2 * \beta_4) + \beta_3$ is `r round(lm.beta(olsRiskFuk)[4]*lm.beta(olsMorePlantsFuk)[1],3) + round(lm.beta(olsMorePlantsFuk)[6],3)`. We use the same process to calculate the effects of ideology and benefits, as well as Fukushima and perceived risks and benefits on support for more nuclear power plants. The effects are shown in @tbl-med.

|   | Indirect Effect | Direct Effect | Total Effect |
|:--------------------------|--------------:|--------------:|--------------:|
| Increasing Conservatism and Nuclear Risks | `r round(lm.beta(olsRiskFuk)[4]*lm.beta(olsMorePlantsFuk)[1],3)` | `r round(lm.beta(olsMorePlantsFuk)[6],3)` | `r round(lm.beta(olsRiskFuk)[4]*lm.beta(olsMorePlantsFuk)[1],3) + round(lm.beta(olsMorePlantsFuk)[6],3)` |
| Increasing Conservatism and Nuclear Benefits | `r round(lm.beta(olsBenefitFuk)[4]*lm.beta(olsMorePlantsFuk)[2],3)` | `r round(lm.beta(olsMorePlantsFuk)[6],3)` | `r round(lm.beta(olsBenefitFuk)[4]*lm.beta(olsMorePlantsFuk)[1],3) + round(lm.beta(olsMorePlantsFuk)[6],3)` |
| Fukushima and Nuclear Risks | `r round(lm.beta(olsRiskFuk)[6]*lm.beta(olsMorePlantsFuk)[1],3)` | `r round(lm.beta(olsMorePlantsFuk)[7],3)` | `r round(lm.beta(olsRiskFuk)[6]*lm.beta(olsMorePlantsFuk)[1],3) + round(lm.beta(olsMorePlantsFuk)[7],3)` |
| Fukushima and Nuclear Benefits | `r round(lm.beta(olsBenefitFuk)[6]*lm.beta(olsMorePlantsFuk)[2],3)` | `r round(lm.beta(olsMorePlantsFuk)[7],3)` | `r round(lm.beta(olsBenefitFuk)[6]*lm.beta(olsMorePlantsFuk)[2],3) + round(lm.beta(olsMorePlantsFuk)[7],3)` |

: The Effects of Political Ideology, Fukushima, and the Risks and Benefits of Nuclear Energy on Support for More Nuclear Plants {#tbl-med}

As shown in @tbl-med, deep core beliefs had larger total effects on support for more nuclear power plants than the Fukushima accident. These results illustrate the influence of deep core beliefs on policy preferences, even when learning occurs regarding policy core beliefs.

## Conclusion

In this chapter, we use public opinion data about nuclear energy taken across nearly two decades to examine the policy learning through information processing model. While scholars of policy learning are generally concerned with learning among elite policy actors, public opinion can be a useful test bed for understanding the mechanics of our learning model, namely, beliefs and belief change. As discussed above, research on belief systems in mass publics suggests that elite-like belief structures are concentrated among the more attentive and sophisticated segment of the public, and restricting our analysis to respondents with an advanced degree is intended to approximate that segment.

The model posits that learning as belief updating is a function of the strength of the information signal relative to the strength of the prior belief, and only when the information signal is sufficiently strong does belief shift. In addition, belief systems are hierarchical and three-tired, with deep core beliefs shaping policy core beliefs and both deep core and policy core beliefs shaping secondary beliefs. While any type of belief can change, secondary beliefs are most likely to change, followed by policy core beliefs, with deep core beliefs least likely to change.

Our findings support the model, as beliefs about nuclear energy changed significantly in response to the Fukushima accident. The Fukushima accident acted as a strong information signal about the risks of nuclear energy. Learning seemed to occur as the policy core beliefs associated with nuclear energy (e.g., the risks and benefits of nuclear energy) changed significantly following Fukushima. Consistent with Fukushima being a signal of risk, risk perceptions increased significantly, while perceptions of the benefits of nuclear energy shifted by a smaller margin. Additionally, support for building more nuclear power plants dropped significantly following Fukushima, as perceptions of risk and benefits shifted. Finally, deep core beliefs acted as filters for information processing, as increasing conservatism was associated with decreasing perceptions of risk, increasing perceptions of benefits, and increased support for building more nuclear power facilities, even while controlling for the impacts of Fukushima. Indeed, @tbl-med illustrates that the deep core belief of ideology had a larger overall effect on nuclear energy beliefs than the Fukushima accident. Thus, even as learning occurred regarding policy core and secondary beliefs, that learning was shaped and likely muted by deep core beliefs.

The model of learning through information processing discussed thus far focuses on beliefs and information processing among policy actors; however, for belief change to lead to policy change, it must be understood within the broader context of policymaking. In the next chapter, we examine how the model fits into the broader policymaking system and how learning can lead to policy change.

## Appendix

Table 3A.1 reports descriptive statistics for the variables used throughout this chapter's analysis, and Table 3A.2 reports the pairwise correlations among the core continuous variables, both restricted to the advanced-degree subsample used in the analyses above.

**Table 3A.1: Descriptive Statistics for Variables Used in the Analysis (Advanced-Degree Subsample)**

```{r appendix descriptives}
#| echo: false
#| warning: false

library(modelsummary)

descVars <- subset(nukeData, grad==1)[, c("nukeRisk","nukeBenefit","morePlants","ideology","age","gender","white","fuk")]
colnames(descVars) <- c("Risks","Benefits","More Plants","Ideology","Age","Gender (male=1)","Race (white=1)","Fukushima")

datasummary_skim(descVars,
  fun_numeric = list(Mean = Mean, SD = SD, Min = Min, Median = Median, Max = Max,
                      "Missing Pct." = PercentMissing))
```

**Table 3A.2: Pairwise Correlations Among Core Continuous Variables (Advanced-Degree Subsample)**

```{r appendix correlations}
#| echo: false
#| warning: false

corrVars <- descVars[, c("Risks","Benefits","More Plants","Ideology","Age")]
datasummary_correlation(corrVars)
```

As shown in Table 3A.1, the analytic sample has some missing data on the risk, benefit, and more-plants scales (4-6%), consistent with standard item non-response on multi-item survey batteries. Table 3A.2 shows that perceived risks and perceived benefits of nuclear energy are only weakly correlated with each other, while support for more nuclear plants is more strongly correlated with both risk and benefit perceptions than with any other variable in the model, consistent with the issue-definition mechanism discussed in the chapter.